---
title: 'Peer-graded Assignment: Course Project 1'
author: "Ngnotue"
date: "02/ 09/ 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Loading and preprocessing the data

### 1. We laod and read the csv file
```{r}
data <- read.csv("activity.csv", header=TRUE)
```
## What is mean total number of steps taken per day?

### 1. We Subset the original data to calculate the total number of steps
```{r}
totalStepsByDay <- aggregate(steps~date, data, sum)
head(totalStepsByDay)
```
### 2. We make a histogram of the total number of steps taken each day
```{r}
hist(totalStepsByDay$steps, col = "red"
     , xlab = "Steps", main = "Total Number of Steps Per Day")
```
### 3. We determine and report the mean and median total number of steps taken per day

#### a. Mean
```{r}
mean1 <- mean(totalStepsByDay$steps)
```
```{r}
mean1
```

#### b. Median
```{r}
median1 <- median(totalStepsByDay$steps)
```
```{r}
median1
```

## What is the average daily activity pattern?

### 1. We subset the original data to calculate the average number of steps taken of the 5-minute interval
```{r}
averageSteps <- aggregate(steps~interval, data, mean)
head(averageSteps)
```

### 2. We make a time series plot
```{r}
with(averageSteps, plot(interval, steps
                        , type="l", main="Average number of steps taken in 5-min interval"
                        , col = "turquoise4")) 
```

#### the maximum data is:
```{r}
max <- max(averageSteps$steps)
max
```

#### The interval contains max data is:
```{r}
subsetData <- subset(averageSteps, steps == max)
subsetData
subsetData$interval
```

## Imputing missing values

### 1. Calculate and report the total number of missing values
```{r}
colSums(is.na(data))
```
### 2. Create a new column "newSteps"
```{r}
data$newSteps <- averageSteps$steps
head(data)
```
### 3. Fill in the missing data

```{r}
data$steps[is.na(data$steps)] <- data$newSteps
```
### 4. Create a new dataset
```{r}
data$newSteps <- NULL
newdata <- data
head(newdata)
```
### 5. Subset in the new data set to calculate the total number of steps per day
```{r}
sumdata <- aggregate(steps~date, newdata, sum)
head(sumdata)
```
### 6. Make a histogram
```{r}
hist(sumdata$steps, col = "blue"
     , xlab = "Steps"
     , main = "Total Number of Steps Per Day (with the missing data filled in)")
```
### 7. Calculate and report the mean and median

#### a. Mean
```{r}
mean2 <- mean(sumdata$steps)
```
```{r}
mean2
```

#### b. Median
```{r}
median2 <- median(sumdata$steps)
```
```{r}
median2
```

#8. Calculate the differences
```{r}
mean2 - mean1
```
```{r}
median2 - median1
```

## Are there differences in activity patterns between weekdays and weekends?
#load the timeDate package
# install.packages("timeDate")
```{r}
library(timeDate)
```
### 1. Create a new column, and use the isWeekday() function to check if the date is weekday or weekend
```{r}
newdata$Weekday <- isWeekday(newdata$date)
head(newdata)
```

### 2. Subset and calculate the average steps for weekday and weekend
```{r}
weekday <- subset(newdata, newdata$Weekday == "TRUE")
weekdayMean <- aggregate(steps ~ interval, data = weekday, mean)
head(weekdayMean)
```
```{r}
weekend <- subset(newdata, newdata$Weekday == "FALSE")
weekendMean <- aggregate(steps ~ interval, data = weekend, mean)
head(weekendMean)
```
### 3. We Make the panel plot to calculate the average number of steps taken for weekday 
```{r}
layout(matrix(c(1,1,2,2), 2, 2, byrow = TRUE))
```
```{r}
plot(weekdayMean$interval, weekdayMean$steps
     , xlab = "interval", ylab = "Number of steps"
     , main ="Weekday", col ="yellow", type="l") 
```
###4. We Make the panel plot to calculate the average number of steps taken for  weekend
```{r}
plot(weekendMean$interval, weekendMean$steps
     , xlab = "interval", ylab = "Number of steps"
     , main ="Weekend", col ="darkgreen", type="l")
```
